Advantages of mixed unitary operators for quantum information processing

### 2018-12-26 decorrelation/mixed_unitaries.tex
pulling over more of the work from the first tex document

\author{Anthony Polloreno}
\email[Email: ]{anthony@rigetti.com}
\affiliation{Rigetti Computing, Berkeley, CA}

@@

In this article we discuss various applications of mixed unitary controls, and show that the advantages of this approach can be made robust to drift in the target gates. In this article, we discuss an optimal control approach to the mixed unitary control design problem.  We apply our methods in simulation where we construct single- and two-qubit mixed unitary controls which are robust to drift and uncertainty in the control parameters. We further present an experimental implementation of single-qubit mixed unitary controls on a superconducting qubit testbed at Rigetti Quantum Computing. Using randomized benchmarking, we are able to show a marked improvement in error rates, as well as a reduced variance in circuit outcome probabilities, indicating a reduction in the coherence of the error.

@@

** The vec representation may be a little more convenient here? The channel that results is:
\begin{align}
	\mathcal{M} 
		=& \sum_i p_i U^*_i \otimes U_i \\
		=& \sum_i p_i \exp(i \theta \hat n \cdot \vec\sigma^*) \otimes \exp(-i \theta \hat n \cdot \vec\sigma) \\
		\simeq & \sum_i p_i (\cos\theta + i \sin\theta H^*)\otimes(\cos\theta -i \sin\theta H) \\
		\simeq & \sum_i p_i (\cos\theta^2 I\otimes I + \sin\theta^2 H^*\otimes H \\ & + i\sin\theta\cos\theta H^*\otimes I
								- i\sin\theta\cos\theta I\otimes H )
\end{align}
In this case, $\sin\theta^2$ is always positive, so we cannot hope to eliminate the second term, but $\sin\theta\cos\theta$ may be positive or negative, and so there is hope that we could possibly combine various implementations to eliminate this term. The result would be a purely stochastic channel.

@@

\end{center}
Make a table of these error channels.

@@

** This example is really useful and should be expanded. We should discuss the various optimization problems, a mixed unitary strategy, the fidelity of the resulting channel.

@@

\note{Anthony mentions the sparsity constraint, as well as the $\ell_1$ penalty, as being described in this section.}

While mixed unitary

@@

@@

\section{Random Gate Synthesis}
\label{sec:random_gate_synthesis}
 In this paper, we use optimal control to generate the one qubit families of controls for MUPs. One could use any of the many available quantum optimal control techniques \cite{Khaneja2005, Caneva2011, Machnes2018}, and for our numerics we chose to use the GRAPE algorithm. First described in \cite{Khaneja2005}, the GRAPE (GRadient Ascent Pulse Engineering) algorithm is a technique for finding piecewise constant control sequences that approximate a desired unitary, $U_T$, given a Hamiltonian with controlled and uncontrolled terms. Defining the uncontrolled Hamiltonian as $H_0$, the control Hamiltonians as $H_{i\neq 0}$, and the \textit{control matrix} $c_{ij}$ as containing the control amplitude associated with the $i^{th}$ time step and the $j^{th}$ Hamiltonian, we can write an the unitary for any timestep of evolution as:
\begin{equation}\label{eq:3}
  U_i = \exp\{-i\Delta t(H_0 + \sum_{j=1}^{n}c_{ij}H_{j}\}
\end{equation}
To measure the simularity of our approxiate unitary $U=\prod_iU_{i=1}^n$ to our target unitary $U_T$, we can define a cost function $J(U) = Tr\{U_T^{\dagger}U\}$.

@@

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{rb_data.pdf}
  \caption{Randomized benchmarking experiments ran using different pulse definitions. The four plots on the left are from the incorrectly calibrated pulse, while the top right is the calibrated pulse, and the bottom right is the BCS.}
  \label{fig:rb}
\end{figure}

% ==============================================================================
% Section: Figures
% ==============================================================================
\section{Figures}
\label{sec:figures}

>>>>

\author{Anthony M. Polloreno}
\email[Email: ]{anthony@rigetti.com}
\affiliation{Rigetti Computing, Berkeley, CA}

@@

In this article we discuss various applications of mixed unitary controls, and show that the advantages of this approach can be made robust to drift in the target gates. We present an optimal control approach to the mixed unitary control design problem. We apply our methods in simulation where we construct single- and two-qubit mixed unitary controls which are robust to drift and uncertainty in the control parameters. We further present an experimental implementation of single-qubit mixed unitary controls on a superconducting qubit testbed at Rigetti Quantum Computing. Using randomized benchmarking, we are able to show a marked improvement in error rates, as well as a reduced variance in circuit outcome probabilities, indicating a reduction in the coherence of the error.

@@

\note{Consider the vec representation} The channel that results is:
\begin{align}
	\mathcal{M} 
		=& \sum_i p_i U^*_i \otimes U_i \\
		=& \sum_i p_i \exp(i \theta \hat n \cdot \vec\sigma^*) \otimes \exp(-i \theta \hat n \cdot \vec\sigma) \\
		\simeq & \sum_i p_i (\cos\theta + i \sin\theta H^*)\otimes(\cos\theta -i \sin\theta H) \\
		\simeq & \sum_i p_i (\cos\theta^2 I\otimes I + \sin\theta^2 H^*\otimes H \\ & + i\sin\theta\cos\theta H^*\otimes I
								- i\sin\theta\cos\theta I\otimes H )
\end{align}
In this case, $\sin\theta^2$ is always positive, so we cannot hope to eliminate the second term, but $\sin\theta\cos\theta$ may be positive or negative, and so there is hope that we could possibly combine various implementations to eliminate this term. The result would be a purely stochastic channel.

@@

\end{center}

@@

\note{This example is really useful and should be expanded. We should discuss the various optimization problems, a mixed unitary strategy, the fidelity of the resulting channel.}

@@

While mixed unitary processes offer significant improvements to gate performance, they fail to take into account the reality that most control electronics experience drift over time scales relevant to QIP performance. Because of this drift, the quality of the MUP will degrade. Thus, we would like to design mized unitary processes that are \textit{robust} to this drift. A mixed unitary process is said to be robust to order $\ell$ if for all $1 \leq j \leq \ell$:
\begin{equation}\label{eq:MUP}
\begin{gathered}
\sum_k\omega_k(\sum_{n=0}^j D^n_k)^j = 0\\
D^n_k = \frac{1}{n!}\frac{\partial^{n}}{\partial\delta_{i_1}...\partial\delta{i_n}}H_k(t,\vec{\delta})|_{\delta=\vec{0}}
\end{gathered}
\end{equation}
These conditions imply that an $\ell^{th}$ order MUP ($\ell$MUP) is insensitive to the $\ell^{th}$ order in drift in $\vec{\delta}$. To see this, write the error on each gate in the MUP as 
\begin{equation}\label{eq:taylor}
\begin{gathered}
U_j(\vec{\delta}) = exp(-i(\tilde{H}_j(\vec{0}) + \frac{\partial}{\partial\delta_i}\tilde{H}_j(d\delta_i)\\ +  \frac{1}{2}\frac{\partial^2}{\partial\delta_i\partial\delta_k} \tilde{H}_j(d\delta_i d\delta_k) + ...)U_T
\end{gathered}
\end{equation}
where $\tilde{H}(\delta)$ generates the error $U_j(\vec{\delta})U_T^{\dagger}$ over variations in $\vec{\delta}$. Then, by Taylor expanding $U_j(\vec{\delta})U_T^{\dagger}$ one finds that the the first $\ell$ derivatives of an $\ell RBC$ will be zero.

Following from Lemma 2 in \cite{Campbell2017}, if $0\in $ Conv$[\{\tilde{H}_i(t, \vec{\delta})\}]$, where Conv is the convex hull of its arguments, then we know that $\vec{\omega}$ exists that quadratically decreases the diamond norm. Then, following from the mixing lemma proven by both Campbell and Hastings\cite{1612.01011}, we can bound the the diamond distance from a balanced control to $U_T$ by $\epsilon^2$. We can relax Equation \ref{eq:MUP}, to be
\begin{equation}\label{eq:RBC-relaxed}
\begin{gathered}
\sum_k\omega_kD^n_k = 0\\
\end{gathered}
\end{equation}
In this context, since we are assuming that the derivatives are all of norm $\epsilon$, the above condition guarantees that the error unitary be no worse than $\sum^n\epsilon^2\delta^n$ near the origin -- all orders can be quadratically supressed. Figure \ref{fig:vectorspace} gives geometric intuition for robustly mixed unitary processes.

@@

% ==============================================================================
% Section: Generationg of Robustly Mixed Unitary Processes
% ==============================================================================
\section{Generation of Robustly Mixed Unitary Processes}
\label{sec:generation}

To generate robustly mixed unitary processes, we can solve the following convex optimization problem:

\begin{equation}\label{eq:minimization}
  \begin{split}
    &\underset{\omega_j\geq0, |\omega|_1=1}{\textbf{minimize}: } ||D_{\ell}^T\omega||\\
    &\textbf{subject to: } \forall i<\ell, \sum \omega_jD_j^i=0\\
  \end{split}
\end{equation}

If the control family is too small a solution won't exist, in particular as in \cite{Campbell2017} we need  $0\in $ Conv$[D_j^i]$.  This can be accomplished by modifying the oracle in Campbell's algorithm to produce robust controls, in particular by updating the GRAPE cost function to include derivative information, one could produce controls with derivatives of order $\epsilon$. In our case, by using the Gaussian weighted cost function, we ensure that the controls generated should have vanishing first derivative.

This minimization problem, while sufficient for quadratically decreasing the diamond norm relative to the \textit{worst} controls in the family, does not preferentially select the controls with the least error. That is to say, as in Section \ref{sec:mixed_unitary_processes}, both \{$U_{+2\epsilon}$, $U_{-2\epsilon}$\} and \{$U_{\epsilon}$, $U_{-\epsilon}$\} minimize this cost function for a 0MUP. To encourage the inclusion of controls with smaller error, we may impose an $\ell_2$ penalty to our cost function by rewriting it as:

\begin{equation}\label{eq:minimization_l2}
  \begin{split}
    &\underset{\omega_j\geq0, |\omega|_1=1}{\textbf{minimize}: } ||D_{\ell}^T\omega||_2 + \eta\sum||D_0^T\omega||_2\\
  \end{split}
\end{equation}
with $\eta \geq 0$.

To make sure the routine remains practical, we would also like to regularize our objective function to enforce sparsity. In this case, lasso regularization \cite{tibshirani1996regression} is insufficient as we already constrain the one norm of the vector we optimize over to be one. However, the problem of enforcing sparsity in such situations has been considered in \cite{NIPS2012_4504} and can be expressed via another convex program that extends Equation \ref{eq:minimization}:
\begin{equation}
\begin{split}
\underset{n\in[N]}{\textbf{minimize}}\{
    &\underset{\omega_j\geq0, |\omega|_1=1, t\geq0}{\textbf{minimize}: } ||D_{\ell}^T\omega|| + t\\
    &\textbf{subject to: } \omega_n > \frac{\lambda}{t}\\
    &\phantom{\textbf{subject to: }} \forall i<\ell, \sum \omega_jD_j^i=0\}
\end{split}
\end{equation} where $\lambda$ is a hyperparameter to be optimized over.

@@

\section{Random Gate Synthesis}
\label{sec:random_gate_synthesis}
 In this article, we use optimal control to generate the one qubit families of controls for MUPs. One could use any of the many available quantum optimal control techniques \cite{Khaneja2005, Caneva2011, Machnes2018}, and for our numerics we chose to use the GRAPE algorithm. First described in \cite{Khaneja2005}, the GRAPE (GRadient Ascent Pulse Engineering) algorithm is a technique for finding piecewise constant control sequences that approximate a desired unitary, $U_T$, given a Hamiltonian with controlled and uncontrolled terms. Defining the uncontrolled Hamiltonian as $H_0$, the control Hamiltonians as $H_{i\neq 0}$, and the \textit{control matrix} $c_{ij}$ as containing the control amplitude associated with the $i^{th}$ time step and the $j^{th}$ Hamiltonian, we can write an the unitary for any timestep of evolution as:
\begin{equation}\label{eq:3}
  U_i = \exp\{-i\Delta t(H_0 + \sum_{j=1}^{n}c_{ij}H_{j}\}
\end{equation}
To measure the simularity of our approxiate unitary $U=\prod_iU_{i=1}^n$ to our target unitary $U_T$, we can define a cost function $J(U) = Tr\{U_T^{\dagger}U\}$.

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=\columnwidth]{rb_data.pdf}
  \caption{Randomized benchmarking experiments ran using different pulse definitions. The four plots on the left are from the incorrectly calibrated pulse, while the top right is the calibrated pulse, and the bottom right is the BCS.}
  \label{fig:rb}
\end{figure}

@@
